Association between trauma severity models and opportunities for improvement:

### 2022-10-13 manuscript.Rmd
made ofi_diagram() a bit better perhaps
some progress on results

---
title: 'Association between trauma severity models and opportunities for improvement:
  A retrospective cohort study'
author: "Victoria Witthoff"
output:
  html_document: default
  pdf_document: default
  word_document: default
bibliography: bibliography.bib
csl: vancouver.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

```{r main, echo=FALSE, message = FALSE}
  source("main.R") 
  exclusion_flowchart(inclusion.counts)
```

@@

Participants
------------

<!-- Cohort study: Give the eligibility criteria, and the sources and
methods of selection of participants. Describe methods of
follow-up. For matched studies, give matching criteria and number of
exposed and unexposed -->

<!-- Case-control study: Give the eligibility criteria, and the
sources and + methods of case ascertainment and control
selection. Give the rationale for the choice of cases and
controls. For matched studies, give matching criteria and the number
of controls per case -->

<!-- Cross-sectional study: Give the eligibility criteria, and the
sources and methods of selection of participants -->

@@

Participants
------------

(a) Report numbers of individuals at each stage of study—eg numbers
potentially eligible, examined for eligibility, confirmed eligible,
included in the study, completing follow-up, and analysed

@@

@@

(a) Give characteristics of study participants (eg demographic,
clinical, social) and information on exposures and potential
confounders

@@

(c) Cohort study — Summarise follow-up time (eg, average and total
amount)

@@

Cohort study — Report numbers of outcome events or summary measures
over time

Case-control study — Report numbers in each exposure category, or
summary measures of exposure

Cross-sectional study — Report numbers of outcome events or summary
measures

@@

(a) Give unadjusted estimates and, if applicable, confounder-adjusted
estimates and their precision (eg, 95% confidence interval). Make
clear which confounders were adjusted for and why they were included

(b) Report category boundaries when continuous variables were
categorized

(c) If relevant, consider translating estimates of relative risk into
absolute risk for a meaningful time period

@@

Report other analyses done—eg analyses of subgroups and interactions,
and sensitivity analyses 
-->

>>>>

---
title: 'Association between trauma severity models and opportunities for improvement:
  A retrospective cohort study'
author: "Victoria Witthoff"
output:
  html_document: default
  pdf_document: default
  word_document: default
bibliography: bibliography.bib
csl: vancouver.csl
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r main, echo=FALSE, message = FALSE}
  source("main.R") 
```

@@

@@

Participants
------------
<!-- Cohort study: Give the eligibility criteria, and the sources and
methods of selection of participants. -->
Participants were patients registered in both the trauma registry and the trauma care quality database at KUH. The eligibility criteria for this study were the criteria for the two databases used in addition to some study-specific inclusion criteria.

### Inclusion criteria for the trauma registry

The trauma registry includes all patients for whom the trauma team was activated after receiving a potentially traumatic injury, regardless of NISS score, and all patients with a NISS score >15, regardless of whether the trauma team was activated. Also included are patients who were transferred to the hospital within 7 days of injury and have a NISS score >15. Patients where the only traumatic injury is a chronic subdural hematoma and patients for whom the trauma team was activated without an underlying traumatic injury are excluded from the register.

### Inclusion criteria for the trauma care quality database

The trauma care quality database contains patients from the trauma registry that were selected for M&M review. At KUH, all fatalities are automatically selected for M&M review. In addition to this, the electronic medical records of all trauma patients are screened by research nursing staff to identify cases that potentially received sub-optimal care. These nurses also use audit filters that automatically highlight cases with abnormal parameters. Cases that two research nurses have evaluated and found to have potential for sub-optimal care are also selected for M&M review. Figure 1 gives an overview of the workflow that leads to determination of OFI.

``` {r, echo = FALSE,} 
  ofi_diagram() ## Only works in HTML
```

### Study-specific exclusion criteria
Patients <15 years of age were excluded as paediatric trauma is a separate clinical entity. Patients that were dead on arrival were also excluded. In addition, this is a complete case study and all patients that lacked parameters necessary to carry out the analysis were excluded.

@@

Participants
------------
<!-- 
(a) Report numbers of individuals at each stage of study—eg numbers
potentially eligible, examined for eligibility, confirmed eligible,
included in the study, completing follow-up, and analysed

@@

-->
A total of `r counts[6,3]` patients were excluded during processing of data. `r counts[2,3]` patients were excluded due to not having a recorded outcome from an M&M conference. `r counts[3,3]` patients were excluded for being under 15 years of age. `r counts[4,3]` patients were excluded due to being dead on arrival at hospital. `r counts[5,3]` patients were excluded due to missing required data. Figure 2 shows the selection process for eligible participants.

```{r, echo = FALSE}
  exclusion_flowchart(counts) ## Only works in HTML
```

@@

<!--
(a) Give characteristics of study participants (eg demographic,
clinical, social) and information on exposures and potential
confounders

@@

(c) Cohort study — Summarise follow-up time (eg, average and total
amount) -->

A total of `r counts[5,2]` participants were identified as eligible for inclusion. Their demographical and clinical characteristics are described in Table 1. `r numbers['ofi.true']` cases had at least one OFI and `r numbers['ofi.false']` cases had no OFI. The mean age for all patients was `r numbers['mean.age']` and `r numbers['pc.male']`% were male.

For patients with at least one OFI, the mean TRISS score was `r numbers['t.ofi.t']` and the mean NORMIT score was `r numbers['n.ofi.t']`. For patients with no OFI, the mean TRISS score was `r numbers['t.ofi.f']` and the mean NORMIT score was `r numbers['n.ofi.f']`.

```{r, echo = FALSE} 
  table.one ## will be improved later
```

@@

<!-- Cohort study — Report numbers of outcome events or summary measures
over time -->

@@

The odds ratio for TRISS was `r t.or["or"]` (95% CI: `r t.or["5 %"]`-`r t.or["95 %"]`). When plotting the ROC, the AUC was `r t.auc` (95% CI: `r t.auc.ci[1]`-`r t.auc.ci[3]`).

@@

<!-- Report other analyses done—eg analyses of subgroups and interactions,
and sensitivity analyses -->
